---
title: "ESM 232 Assignment 6"
author: "Madeline Oliver, Ruben Sanchez, Atefeh Mohseni"
date: "5/5/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)

library(tidyverse)
library(deSolve)
library(sensitivity)
library(tinytex)
```

# Model of forest growth (where forest size in measured in units of carbon (C))

* dC/dt  = r*C for forests where carbon is below a canopy closure threshold 

* dC/dt = g  for forests where carbon is  at or above the threshold canopy closure

* dC/dt = 0 once a carrying capacity (K) is reached.

The size of the forest (C), Canopy closure threshold and carrying capacity are all in units of carbon 

**Canopy closure threshold:** the size of the forest at which growth rates change from exponential to linear

**r:** early exponential growth rate 

**g:** linear growth rate once canopy closure has been reached

# Run model for 300 years (using the ODE solver) starting with an initial forest size of 10 kg/C and using the following parameters:

* canopy closure threshold (thresh)  = 50 kgC 

* K = 250 kg C (carrying capacity) 

* r=  0.01 (exponential growth rate before before canopy closure)

* g = 2 kg/year (linear growth rate after canopy closure)


```{r}
# source model
source("forest_growth.R")

# set parameters
years = seq(from=1, to=300, by=1)
initial_c = 10
thresh = 50
K = 250
r = 0.01
g = 2

# create parameter list
parms = list(thresh = thresh, K=K, r=r, g=g)

# run ODE 
result = ode(y=initial_c, 
              time=years, 
              func=forest_growth, 
              parms=parms)

result = as.data.frame(result)

colnames(result)=c("year","forest_size")
head(result)

# graph result
ggplot(result, aes(year,forest_size))+
  geom_point()+
  labs(y="Forest Carbon", "Years")

```

# Sobol sensitivity analysis 

Our sobol sensitivity analysis explores how estimated max and mean forest size (C) varies with the: 

* pre canopy closure growth rate(C) 

* post-canopy closure growth rate (g) 

* canopy closure thresholds (thresh)

* carrying capacity (K)

Assume all parameters are normally distributed with means as given in previous section and standard deviation of 10% of mean value

```{r}
# set number of paramenters
np = 1000

# create two sets of sample parameters
thresh = rnorm(mean = 50, sd = .01*50, n = np)
K = rnorm(mean = 250, sd = .01*250, n = np) 
r = rnorm(mean = 0.01, sd = .01*.01, n = np)
g = rnorm(mean = 2, sd = .01*2, n = np)
x1 = cbind.data.frame(thresh=thresh, 
                      K=K, 
                      r=r, 
                      g=g)

# repeat to get second set of samples
thresh = rnorm(mean = 50, sd = .01*50, n = np)
K = rnorm(mean = 250, sd = .01*250, n = np) 
r = rnorm(mean = 0.01, sd = .01*.01, n = np)
g = rnorm(mean = 2, sd = .01*2, n = np)
x2 = cbind.data.frame(thresh=thresh, 
                      K=K, 
                      r=r, 
                      g=g)

# create sobol object and get sets of parameters for running the model

sens_c = soboljansen(model = NULL, x1, x2, nboot = 300)

# our parameter sets are:
head(sens_c$X)

# get results for 300 years, evaluating every year
simtimes = seq(from=1, to=300)
parms = list(thresh=sens_c$X$thresh[1], K=sens_c$X$K[1], r=sens_c$X$r[1], g=sens_c$X$g[1])
result = ode(y=initial_c, times=simtimes, func=forest_growth, parms=parms)

head(result)
colnames(result)=c("time","forest_size")
# turn it into a data frame
result = as.data.frame(result)


```


```{r}
# compute metrics using a function 

compute_metrics = function(result) {
  max_c = max(result$forest_size)
  mean_c = mean(result$forest_size)
  idx = which.max(result$forest_size)
  maxyear = result$time[idx]
return(list(max_c=max_c, mean_c=mean_c, maxyear))}

compute_metrics(result)
```


```{r}
# apply ode and this function for all our parameters by creating a wrapper function that runs solver and computes metrics and returns results for each parameter

c_wrapper = function(time, initial_c, r, g, thresh, K, func){ 
    parms = list(thresh=thresh, 
                 K=K,
                 r=r,
                 g=g)
    result = ode(y=initial_c, 
                 times=time, 
                 func=forest_growth, 
                 parms=parms) 
    colnames(result)=c("year","forest_size")
    
  # get metrics
  metrics=compute_metrics(as.data.frame(result))
 
   return(metrics)
}
```


```{r}
# Use pmap

allresults = sens_c$X %>%
  pmap(c_wrapper, 
       time=time, 
       initial_c=initial_c, 
       r=r, 
       g=g, 
       thresh=thresh,
       K=K, 
       func=forest_growth)

```


```{r}
# extract metrics of interest - max and mean forest size

# extract max forest size
max_c = max(result$forest_size)
max_c

# extract mean forest size
mean_c = mean(result$forest_size)
mean_c

# years required to get to the max forest size


```



